---
title: "Pretty I/O to the console"
author: "Giulio Caravagna"
date: "29/04/2019"
output:
    github_document
---

The **pretty I/O** R package `pio` provides functions to layout coloured outputs to console, in a neat way. If you work constantly with R scripts that generate several outputs, a tidy organisation of your outputs makes it easier to spot relevant information on the fly.

`pio` relies on [crayon](https://github.com/r-lib/crayon), which provides some simple functions to add colours to terminals. This vignette is generated using the [fansi](https://github.com/brodieG/fansi) package that renders ANSI modified strings from `crayon` into HTML code.

```{r, comment="", results="asis"}
# The hook for this vignette, via fansi
old.hooks <- fansi::set_knit_hooks(knitr::knit_hooks)

# Activate crayon and pio
options(crayon.enabled = TRUE)
require(pio)
```


## Example: header and parameters

A header is a one-liner information used to provide some title; I think of that as something used to communicate the name of tool, model, algorithm etc. 

It's simpler form is just some lines of text.
```{r, echo = TRUE, comment=""}
# Simplest possible
pioHdr("My new cool tool")

# Ellipsis are supported
pioHdr("My new cool tool", "- COOLTOOL version ", 12345)
```

A header with also some parameters, if the parameters are not part of a named vector, automatic names are generated. Instead, if names are available, they are used. 

In general all parameters are tokenized via functions `pioTkn` and `pioStr`, which we discuss later.
```{r, echo = TRUE, comment=""}
# Parameters
x = 12345
y = "yes"
z = NA # z will be tokenized because it is a special value (NA)

pioHdr("My new cool function", toPrint = c(x, y, z))

# With names
params = c(x, y, z)
names(params) = c("Input = ", "Test = ", "Plot = ")

pioHdr("My new cool function", toPrint = params)
```

Parameters can be laid out using extra formats like an intendation (or any othe prefix), that is added to all values in `toPrint`. This is also done using `pioStr`, which we discuss later.
```{r, echo = TRUE, comment=""}
pioHdr("My new cool function", toPrint = params, prefix = '\t- ')
```


## Example: titles



We can easily change colours -- black text on orange background
```{r, echo = TRUE, comment=""}
colorscheme_orange_black()

pioHdr("My new cool function", toPrint = params)
```


