---
title: "Pretty I/O to the console"
institute: "Institute for Cancer Research"
author: "Giulio Caravagna"
date: "29/04/2019"
output: 
  # rmarkdown::github_document
  html_document
---

The **pretty I/O** R package `pio` provides functions to print coloured outputs in an organized way to console. This allows organizing outputs in a tidy way, making it easier to spot information while interacting with a tool.

`pio` provides a few types of functions that display:

* a *header*, that is meant to display a summary string (like a program name, a function name etc.);
* a *title*, that is meant to display a string surrounded by some separators (like a title for a task within a multi-step function);
* a *string* in the format `Variable = Value`, meant to display a pair (descriptor plus value).

Other functions and simple ways to manipulate color schema are available and shown in this vignette.

**Note.** `pio` relies on [crayon](https://github.com/r-lib/crayon), which provides core functions to add colours to terminals. This vignette is generated using the [fansi](https://github.com/brodieG/fansi) package that renders ANSI modified strings from `crayon` into HTML code. However, some of the generated layout still renders poorly in HTML, but try the package yourself to see how the layout looks like.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, comment="", results="asis", echo=TRUE, message=FALSE}
# The hook for this vignette, via fansi
old.hooks <- fansi::set_knit_hooks(knitr::knit_hooks)

# Activate crayon and pio
options(crayon.enabled = TRUE)
require(pio)
```


## Header 

A header is a one-line information that can be used to communicate the name of a tool, model, algorithm etc. 

It's simpler form is just some lines of text, which in the baseline setting is white on blue background. The function accepts an ellipsis to allow easy concatenation of multiple values.
```{r, echo = TRUE, comment=""}
version = '1.0'
tool_name = 'COOLTOOL'

pioHdr("My new", tool_name, version)
```

The header function can accepts also parameters, and use other `pio` functions to print them (`pioTkn` and `pioStr`, see below). This helps using `pioHdr` in the beginning of a compplex function call, where we want to summarize some behaviour to the user. 
Notice that, if the parameters have no names associated, automatic names are generated. 

```{r, echo = TRUE, comment=""}
# Parameters
x = 12345
y = TRUE
z = NA  # Notice this how it gets printed

# Vector (without names)
params = c(x, y, z)

pioHdr("My new", tool_name, version, 
       toPrint = params)

# Add names, and re-print
names(params) = c("Input = ", "Test = ", "Plot = ")

pioHdr("My new", tool_name, version, 
       toPrint = params)
```

The list of parameters can be displaied with some extra prefix, (e.g., an intendation, a separator etc.), that is added to all values in `toPrint` using the `pioStr` function. For instance, if we want to add a tabulation to arrange all parameters we can use `prefix = '\t- '`.

## Titles

A title is just a piece of text surrounded by some nice separator. In its simpler form is can contain one or more strings (like the header, via an ellipsis).

```{r, echo = TRUE, comment=""}
# A title with variables
x = 12345
y = "myFun"

pioTit('We now call', y, 'with', x, "as input")
```

This function accepts also a `prefix` argument that can ben used to append some string
before all strings print by `pioTit`.

```{r, echo = TRUE, comment=""}
# A title with escape chars in the prefix
pioTit('We now call', y, 'with', x, "as input",
       prefix = '>--\\/\\/\\/-->>\t')
```

The separator can also be customized.
```{r, echo = TRUE, comment=""}
pioTit('We now call', y, 'with', x, "as input", 
       separator_motif = '~')
```

# Strings

Function `pioStr` can print a pair of values which are thought to represent a description, followed its a value. For instance `P-value = 0.05` could be print by passing the string `"P-value ="` together with `0.05` as arguments to `pioStr`. Notice that the value content is tokenized by function `pioTkn` (see below).

```{r, echo = TRUE, comment=""}
# Simple call
pioStr('N = ', 12345)

# Observe the tokenization
pioStr('A & B = ', TRUE, 'while the result is', NULL)
```

This function accepts both a `prefix` and a `suffix` argument.

```{r, echo = TRUE, comment=""}
x = 1:5

invisible(sapply(x, pioStr,  str = 'N = ', prefix = '\t- ', suffix = '\n'))
```


# Color schema

`pio` provides a default color schema `colorscheme_default`, which is a combination of blue, cyan and white colours from `crayon`. The colours are essentially controlled via some global variables named `pio.XXX`, and there are options to control pretty much the background and foreground colour of each `pio` function.

A colour scheme is defined as in this function
```{r, echo = TRUE, comment=""}
print(colorscheme_default)
```

If you want to change colour scheme you just need to call the appropriate function. There are 2 extra color schema available in the package, `colorscheme_orange_black` which combines yellow and black colours, and `colorscheme_purple_ivory` which combines violet and white colours.

```{r, echo = TRUE, comment=""}
colorscheme_orange_black()

pioHdr("My function name")
pioTit("A long title for my computation")
pioStr("Variable X = ", TRUE)
```


